A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-07-09 papers/earths-future/submission.tex
catalina comments

We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
\citeA{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
There are parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which the data are assumed to come from a ``true'' model which is a member of the finite set of candidate models available to the analyst \cite{bernardo_bayesian:1994}.
Commonly used classical methods for model selection implicitly assume $\mathcal{M}$-closed case and are defined by their ability to asymptotically identify the true model from a set of candidate models.
By contrast, model selection in the $\mathcal{M}$-open case emphasizes a subjective view of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model (in the analyst's subjective judgement) is reached \cite{gelman_workflow:2020}.

@@

To illustrate the proposed decision analytic framework, we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
Following the approach outlined in \citeA{zarekarizi_suboptimal:2020}, we focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and approximate other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
We use the notation developed in the previous section to describe the case study.
Specifically,
\begin{enumerate}
  \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$.
  \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=70$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
  \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
\end{enumerate}
In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).

@@

We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citeA{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
For a discussion of these model outputs we refer the reader to \citeA{ruckert_coastal:2019}.

@@

This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under \gls{noaa} grant NA16OAR4310179 and through the Penn State Initiative for Resilient Communities (PSIRC) by a Strategic Plan seed grant from the Penn State Office of the Provost, with co-support from the Center for Climate Risk Management (CLIMA), the Rock Ethics Institute, Penn State Law, and the Hamer Center for Community Design.
JDG thanks Rice University for support.
KK thanks Dartmouth College for support.
The authors thank
Tor Erlend Fjelde,
Vivek Srikrishnan,
Adam Pollack,
and Sitara Baboolal
for helpful comments.

>>>>

We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
\citeA{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
There are parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which the data are assumed to come from a ``true'' model which is a member of the finite set of candidate models available to the analyst \cite{bernardo_bayesian:1994}.
Commonly used classical methods for model selection implicitly assume $\mathcal{M}$-closed case and are defined by their ability to asymptotically identify the true model from a set of candidate models.
By contrast, model selection in the $\mathcal{M}$-open case emphasizes a subjective view of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model (in the analyst's subjective judgement) is reached \cite{gelman_workflow:2020}.
Thus, where classical statistics, particularly as applied to environmental systems, has often emphasized objectivity, we draw from this Bayesian literature that  emphasizes a focus on building transparency in the presentation, analysis, and critique of models and results.

@@

To illustrate the proposed decision analytic framework, we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
Following the approach outlined in \citeA{zarekarizi_suboptimal:2020}, we focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and approximate other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
We use the notation developed in the previous section to describe the case study.
Specifically,
\begin{enumerate}
  \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$.
  \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=70$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
  \item The system model $f$ quantifies up-front costs (the cost of elevating) and lifetime expected damages (the structural cost of experiencing floods), given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge. We elaborate upon these metrics in \cref{sec:case-metrics}.
\end{enumerate}
In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).

@@

We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citeA{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
These four models represent physical processes, particularly of ice sheet dynamics, in different ways.
The differences in these models leading to diverging estimates of how \gls{msl} responds to temperature.
For a discussion of these model outputs we refer the reader to \citeA{ruckert_coastal:2019}.

@@

This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under \gls{noaa} grant NA16OAR4310179 and through the Penn State Initiative for Resilient Communities (PSIRC) by a Strategic Plan seed grant from the Penn State Office of the Provost, with co-support from the Center for Climate Risk Management (CLIMA), the Rock Ethics Institute, Penn State Law, and the Hamer Center for Community Design.
JDG thanks Rice University for support.
KK thanks Dartmouth College for support.
The authors thank
Sitara Baboolal,
Tor Erlend Fjelde,
Catalina González-Dueñas,
Adam Pollack,
and
Vivek Srikrishnan
for helpful comments.
